Predicting the PCG PSeudo-Random Number Generator In Practice

### 2020-05-13 article/main.tex
lattice+partie de preuve facile

@@

>>>>

\section{lattice geometric progression}

One of the key point of the algorithm is, given a modulus \(M\) and a given common ration \(\alpha\), to find a modular geometric progression knowing only the most significant bits of its \(n\) first terms.

We call \(L(\alpha,M,n)\) the set of n-tuples \(U = (u_0,\dots u_{n-1})\) satisfying :

\[u_{i} = \alpha u_{i-1} \mod M \] for all \(i\) in \(\{1,\dots,n-1\}\).

We easily notice that, if \(U\) and \(V\) are in \(L(\alpha,M,n)\), so are \(-U\) and \(U+V\), hence \(L(\alpha,M,n)\) is a lattice.

The lattice \(L(\alpha,M,n)\) is spawned by the lines of the following matrix :

\[G =\begin{pmatrix}
1&\alpha&\alpha^2&\dots&\alpha^{n-1}\\
0&M&0&\dots&0\\
0&0&M&\dots&0\\
\dots&\dots&\dots&\dots&\dots\\
0&0&0&\dots&M\\
\end{pmatrix}\]

We want to find an element \(U\) in the lattice only knowing the most significant bits of each of its coordinates. We decompose \(U = U_1 + U_2\) where \(U_1\) contains only the upper bits and \(U2\) the lowers one. There is no reason for \(U_1\) to be in the lattice, but as \(U_2\) is small, we know \(U_1\) is \emph{close} to \(U\). So, to find \(U\), we will solve CVP(\(L(\alpha,M)\),\(U_1\)).

@@

\section{babairounding}

We call \(S't\) the truncated vector : \(S'[\ell,64+\ell]\).

We consider the case were we successfully guessed the 6 upper bits of each \(S_i\) and the \(\ell\) lower bits of \(S_0\). Then we know the \(6+\ell\) upper bits of \(S't\).

(à détailer ?)

We had define \(S'\) such that \(S'\) is in the lattice \(L(A,2^{128},n)\) and its \(\ell \) lower bits are 0. Then \(S't\) is in \(L(1,2^{64},n)\).

If we consider the vector \(T'\) computed in step (numerodelastepenquestion), we notice it is close to \(S't\) in the sense of they have exactly the same \(6+\ell\) upper bits. So we want to recover \(S't\) by solving an approach \(CVP(L(A,2^{64},n),T')\).

Now we need to prove that the \(pwet\) obtained with the Babai rounding is indeed \(S't\).

We denote \(r = rounding(YG^{-1}) \).

\begin{align*}
\lVert S't - pwet \rVert &= \lVert S't - rG \rVert \\
&= \lVert S't - (r-YG^{-1} + YG^{-1})G \rVert\\
&= \lVert S't - Y + (r-YG^{-1})G \rVert\\
&\leqslant \lVert S't - Y \rVert + \lVert(r-YG^{-1})\rVert \times \lVert G\rVert\\	
\end{align*}

By definition, \(r\) is the closest integer vector to \(YG^{-1}\). But as \(S't\) is an element of the lattice, \(S'tG^{-1}\) is an integer vector. Thus \(r-YG^{-1}\) is smaller than \(S'tG^{-1}-YG^{-1}\).

Hence :
\begin{align*}
\lVert S't - pwet \rVert &\leqslant \lVert S't - Y \rVert + \lVert(r-YG^{-1})\rVert \times \lVert G\rVert\\	
&\leqslant \lVert S't - Y \rVert + \lVert(S'tG^{-1}-YG^{-1})\rVert \times \lVert G\rVert\\	
&\leqslant \lVert S't - Y \rVert + \lVert(S't-Y)\rVert \times \lVert G^{-1} \rVert  \times \lVert G\rVert\\
& 	\leqslant \lVert S't - Y \rVert \times (1 +\lVert G^{-1} \rVert  \times \lVert G\rVert )\\
\end{align*}

But, by construction of \(Y\), we have that \(\lVert S't - Y \rVert \leqslant \sqrt{n2^{2(58-l)}} \). So

\[ \lVert S't - pwet \rVert \leqslant \sqrt{n2^{2(58-l)}}\times (1 +\lVert G^{-1} \rVert  \times \lVert G\rVert )\]\\

Now all we need to do is find sets of parameters \((n,l)\) such that the following inequality is satisfied: 
\[\sqrt{n2^{2(58-\ell)}} \times (1 +\lVert G^{-1} \rVert  \times \lVert G\rVert) \leqslant \lVert SVP(L(A,2^{64},n))\rVert \].

We notice that only the left term of the inequality depends on \(\ell\).

Hence, some sets of parameters that satisfies this inequality are :
\begin{itemize}
	\item \(n = 3\) and \(\ell \geqslant 19\)
	\item \(n = 4\) and \(\ell \geqslant 13\)
	\item \(n = 5\) and \(\ell \geqslant 11\)
	\item \(n = 6\) and \(\ell \geqslant 8\)
\end{itemize}
